---
title: "Data Mining Thredds"
author: "Janelle Christensen"
date: "8/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# ----------
# purpose: data mining the thredds website
# what will happen: download data from the thredds website
# historical daily, historical monthly, future daily (all models), future monthly (all models)
# adjust as needed
# ----------

library(data.table) 
library(here)
#library(plyr) 
# be careful with this, it causes issues for dplyr::group_by
# I didn't try, but the internet says if you load it before tidyverse, that issue goes away
library(tidyverse)
library(beepr)
library(lubridate)
library(directlabels)
library(data.table)
library(ggbeeswarm)
library(gghalves)
library(directlabels)
library(ggrepel)

```

Quick note: this code runs off the most recent version of R. Additionally, this code requires use of project directories. File organization is key here, there are many articles on how to best organize your files, why projects (rather than using setwd()) are crucial to reproducible data, and why we should all be using them. I won't go into that here, but [this blog](https://martinctc.github.io/blog/rstudio-projects-and-working-directories-a-beginner's-guide/) has a good summary. If you need help understanding how to set up a directory in R, [this website](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects) explains it pretty well.


### To update R:

#### Enter the following code for windows:

1. install.packages("installr")
2. library(installr)
3. updateR()

[More information here] (https://www.r-statistics.com/2015/06/a-step-by-step-screenshots-tutorial-for-upgrading-r-on-windows/#:~:text=If%20you%20are%20running%20R,installr%20updateR()%20%23%20updating%20R.)


#### Enter the following code for mac

1. install.packages('devtools') assuming it is not already installed
2. library(devtools)
3. install_github('andreacirilloac/updateR')
4. library(updateR)
5. updateR(admin_password = 'Admin user password')

[More information here] (http://www.andreacirillo.com/2018/03/10/updater-package-update-r-version-with-a-function-on-mac-osx/)


R may tell you to open RGui

[More information about RGui here] (https://www.dummies.com/programming/r/how-to-navigate-rgui/)

# Set up for the site

Make sure to run this code chunk before running any other code chunks.

```{r}
#initialize with time range of interest and location, these don't get overwritten in loops 
# set up for the site, dates, models, and past data type

site = "little_saddle_mtn" 

lat = 44.702

lon = -110.018

start_past = 1980

end_past = 2019

start_future = 2020

end_future = 2099

model_bc = "inmcm4" #model for "best case" scenario

#options: "inmcm4","NorESM1-M", "MRI-CGCM3", "MIROC5", "IPSL-CM5A-LR", "HadGEM2-CC365", "GFDL-ESM2G", "CanESM2", "CCSM4","CSIRO-Mk3-6-0","CNRM-CM5","BNU-ESM" 

model_bc_rcp = "rcp45" #this will be subbed in the HTML
#options are "rcp85", "rcp45"

model_wc = "HadGEM2-CC365" #model for "worst case" scenario

#options: "inmcm4","NorESM1-M", "MRI-CGCM3", "MIROC5", "IPSL-CM5A-LR", "HadGEM2-CC365", "GFDL-ESM2G", "CanESM2", "CCSM4","CSIRO-Mk3-6-0","CNRM-CM5","BNU-ESM" 

model_wc_rcp = "rcp85" #this will be subbed in the HTML
#options are "rcp85", "rcp45"

past_data = "gridmet"

#options: "gidmet", "daymet"
# if you want to use data that will work with the future models, gridmet is a better choice. It doesn't require adjusting for any biases
```


# Before running the code

Make sure you have created the folders as necessary in your wd. If you don't want to save the figures, then just comment out ggsave inside of the functions.

Run the following code to create all folders in your project folder that will be needed to run this code:

```{r}
# create subfolders for data 

here::here(dir.create(site))

here::here(dir.create(paste(site, 
                            "raw_data",
                            sep = "/")))

here::here(dir.create(paste(site,
                            "raw_data", 
                            paste("lat",lat,"lon",lon,sep = "_"), 
                            sep = "/")))

for (i in c("historical_daily", "historical_monthly", "future_daily", "future_monthly")){


here::here(dir.create(paste(site, 
                            "raw_data",
                            paste("lat",lat,"lon",lon,sep = "_"),
                            i,
                            sep = "/")))

}
                             
```


# Historical Data

Be sure of what the intended use is for your historical data. If comparing it to future data, then you'll want to use the gridmet data.


### Daily Historical Data

```{r}

# ----------
# DAILY HISTORICAL LOOP
# ----------

# set time variables for historical data 

start = start_past

end = end_past

#initialize some variables to hold data 

holder<-NULL 

mydat<-NULL 

mydat2<-NULL 

mydat3<-NULL 



#This is the loop that runs across all the climate variables you want to download 

# "agdd", fix this - return when AGDD is fixed on the website
## make sure the order in colnames at end of loop matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)

for(climvar in c("soil_water", "runoff", "rain", "accumswe", "PET", "Deficit", "AET")){ 
  
  print(paste("downloading",climvar, "daily historical", past_data))#this is a progress update as the loop runs 
  
  
  for(yr in c(start:end)){# for each climate variable loop across the year.  allows users to select range if not interested in all data 
    
    #Specify URL where are stored on cloud 
    
    #access daily data
    
    # 1980 has only 30 days in December, all other years have 31
    
    
   if(past_data == "gridmet"){  #if statement for gridmet data url
     
     leap <- lubridate::leap_year(yr) #gridmet data treats leap years differently
      if(leap == TRUE){
        enddate <- paste(yr,"12-31", sep = "-")
      } else {
        enddate <- paste(yr + 1, "01-01", sep = "-")
      }
          
          data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/v2_historical/gridmet_v_1_5_historical/V_1_5_",yr,"_gridmet_historical_",climvar,".nc4?var=",climvar,"&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-01T12%3A00%3A00Z&time_end=",enddate,"T12%3A00%3A00Z&accept=csv",sep ="")

   } #end if statement for gridmet data url
    
    if(past_data == "daymet"){ #if statement for daymet url
          
          if(yr == 1980){#daymet data has every year ending in 31 other than 1980
      day <- 30
    } else {
      day <- 31
    }
          
    data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/v2_historical/daily/v2_",yr,"_",climvar,".nc4?var=",climvar,"&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-01T12%3A00%3A00Z&time_end=",yr,"-12-",day,"T12%3A00%3A00Z&accept=csv",sep ="")
    
    } #end if statement for dayment data url
    
    holder <-data.frame(fread(data_url, verbose = FALSE, showProgress = FALSE))#temporary holder for subsets downloaded from cloud 
    
    mydat<-rbind(mydat, holder)#file that grows by adding each chunk downloaded from cloud 
    
  }#end loop across years 
  
  
  mydat2<-cbind(mydat2, mydat[,4])#append just the water balance data from each downloaded chunk
  
  date <- mydat$time
  
  mydat<-NULL#reset this variable so it can accodomate a new column name given the new water balance variable it's extracting at each loop cycle
  
  
}#end loop across climate variables 

beep(10)


mydat3<-cbind(date,holder[,2:3],mydat2)#join the data with metadat including date, lat, long 

# "agdd_daily", fix this - return when agdd is back on the website
## make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)

colnames(mydat3)[]<-c("date", "lat","lon","soil_water_daily", "runoff_daily", "rain_daily", "accumswe_daily", "pet_daily", "deficit_daily", "aet_daily") 

if(past_data == "gridmet"){
  
   divide.by.10 <- function(x, na.rm = FALSE) {
    x/10
  }
  
  mydat4 <- mydat3 %>% 
    mutate(lat = as.character(lat),
           lon = as.character(lon)) %>% 
    mutate_if(is.numeric, divide.by.10)
  
  write_csv(mydat4, here::here(site,
                             "raw_data",
                             paste("lat",lat,"lon",lon, sep = "_"),
                             "historical_daily",
                             paste(site,"_daily_", past_data,".csv", sep = ""))) #default is a space, sep = "" removes space
  
}

if(past_data == "daymet"){

write_csv(mydat3, here::here(site,
                             "raw_data",
                             paste("lat",lat,"lon",lon, sep = "_"),
                             "historical_daily",
                             paste(site,"_daily_", past_data,".csv", sep = ""))) #default is a space, sep = "" removes space
  
}


```

### Monthly Historical Data

```{r}

# ------------
# MONTHLY HISTORICAL LOOP
# ------------

# set time variables for historical data 

start = start_past

end = end_past

#initialize some variables to hold data 

holder <- NULL 

mydat <- NULL 

mydat2 <- NULL 

mydat3 <- NULL 



#This is the loop that runs across all the climate variables you want to download 

#"agdd_monthly", fix this - return when agdd is back on the website
## make sure the order in colnames at the end of the loop matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)

for(climvar in c("soil_water_monthly", "runoff_monthly", "rain_monthly", "accumswe_monthly", "PET_monthly", "Deficit_monthly", "AET_monthly")){ 
  
  print(paste("downloading",climvar, past_data, "historical"))#this is a progress update as the loop runs 
  
  
  #the url uses two variable names for same variable in monthly download string like "PET_monthly" for long name and "pet" for short name 
  
  #the short name is sometimes a different case than the long variable name, so make it lowercase so that 
  
  #url string will work with the format of data stored on the cloud 
  
  varshort<-unlist(strsplit(climvar,"_m"));varshort 
  
  vshort<-varshort[[1]];vshort 
  
  if(vshort=="PET"){ 
    
    vshort<-"pet" 
    
  };vshort 
  
  
  
  if(vshort=="Deficit"){#change case of short name from upper to lower case 
    
    vshort<-"deficit" 
    
  };vshort 
  
  
  if(vshort=="AET"){ 
    
    vshort<-"aet" 
    
  };vshort 
  
  
  
  for(yr in c(start:end)){# for each climate variable loop across the year.  allows users to select range if not interested in all data 
    
    #Specify URL where are stored on cloud 
    
    #single variable single year monthly time series with variable and year specified by user 
    
    if(past_data == "gridmet"){  #if statement for gridmet data url
      
      leap <- lubridate::leap_year(yr) 
#leap years act differently with this monthly future data
#need to correct for that
    if(leap == TRUE){
      day <- 16
    } else {
      day <- 17
    }
    
    data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/v2_historical/gridmet_v_1_5_historical/V_1_5_",yr,"_gridmet_historical_",climvar,".nc4?var=",vshort,"&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-16T05%3A14%3A31.916Z&time_end=",yr,"-12-",day,"T00%3A34%3A14.059Z&accept=csv_file",sep ="") 
    
    } #end if statement for gridmet data url
    
    if(past_data == "daymet"){  #if statement for daymet data url
    data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/v2_historical/monthly/v2_",yr,"_",climvar,".nc4?var=",vshort,"&latitude=",lat,"&longitude=",lon,"&time_start=1980-01-16T05%3A14%3A31.916Z&time_end=1980-12-16T00%3A34%3A14.059Z&accept=csv_file",sep ="") 
    } #end if statement for daymet data url
    
    
    holder <-data.frame(fread(data_url, 
                              verbose=FALSE, 
                              showProgress = FALSE,
    ))#temporary holder for subsets downloaded from cloud 
    
    mydat<-rbind(mydat, holder)#file that grows by adding each chunk downloaded from cloud 
    
  }#end loop across years 
  
  mydat2<-cbind(mydat2,mydat[,4])#append just the water balance data from each downloaded chunk 
  
  date <- mydat$time # date will download wrong
  
  mydat<-NULL#reset this variable so it can accodomate a new column name given the new water balance variable it's extracting at each loop cycle
  
  
}#end loop across climate variables 


beep(10)


mydat3<-cbind(date,holder[,2:3],mydat2)#join the data with meta data including date, lat, long 

# "agdd_monthly", fix this - return when agdd is back on the website
## make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)

colnames(mydat3)[]<-c("date", "lat","lon","soil_water_monthly", "runoff_monthly", "rain_monthly", "accumswe_monthly", "pet_monthly", "deficit_monthly", "aet_monthly") 


if(past_data == "gridmet"){
  
   divide.by.10 <- function(x, na.rm = FALSE) {
    x/10
  }
  
  mydat4 <- mydat3 %>% 
    mutate(lat = as.character(lat),
           lon = as.character(lon)) %>% 
    mutate_if(is.numeric, divide.by.10)
  
  write_csv(mydat4, here::here(site,
                               "raw_data",
                               paste("lat",lat,"lon",lon, sep = "_"),
                               "historical_monthly",
                               paste(site,"_monthly_", past_data, ".csv", sep = "")))
}


if(past_data == "daymet") {#dates for daymet data are wrong, they are all 1980

span<-end-start;span#of years downloaded 

year<-rep(start:end,each=12);year 

month<-rep(1:12,span+1);month 

length(year); length(month); nrow(mydat3) 

names(mydat3) 

mydat4<-cbind(year,month,mydat3[,c(2:10)]);head(mydat4)#drop bogus dates and add year and month
#fix this - with agdd back in it should be [,c(2:11)]


write_csv(mydat4, here::here(site,
                             "raw_data",
                             paste("lat",lat,"lon",lon, sep = "_"),
                             "historical_monthly",
                             paste(site,"_monthly_", past_data, ".csv", sep = "")))
} #end if statement for daymet date data
#default is a space, sep = "" removes space

# here::here allows you to choose a folder in your working directory to save to
# here::here("folder I want to save to (this has to be made in your working directory before you save it)", "subfolder", "subfolder ( subfolders can be as numerous you want them to be, always separated by a commma)", "last thing in the list is the name of the object I'm saving")

```

# Future Data

## Future Data mixed models

### Daily future mixed models

This data will download the futures of specified RCPs and models into a single csv. It will mix the model types as specified (i.e. inmcm4 4.5 and CCSM4 8.5 future scenarios) and label the data as bc (best case) and wc (worst case)

```{r}


start = start_future

end = end_future 

  
  holder <- NULL 
  
  mydat <- NULL 
  
  mydat2 <- NULL 
  
  mydat3 <- NULL 
  
  #This is the loop that runs across all the climate variables you want to download for worst case scenario
  
  #"agdd", fix this not a part of the website 8-20-20
  
  
  for(climvar in c("soil_water", "runoff", "rain","accumswe", "PET", "Deficit", "AET")){ #start loop for 8.5
    
    
    
    print(paste("downloading",climvar, "daily", model_wc, model_wc_rcp))#this is a progress update as the loop runs 
    
    
    for(yr in c(start:end)){# for each climate variable loop across the year.  allows users to select range if not interested in all data 
      
      
      #Specify URL where are stored on cloud 
      # leap years have end date set at yr + 1 01-01
      # set if else for future data
      
      leap <- lubridate::leap_year(yr) 
      if(leap == TRUE){
        enddate <- paste(yr,"12-31", sep = "-")
      } else {
        enddate <- paste(yr + 1, "01-01", sep = "-")
      }
      
      data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/gcm/",model_wc_rcp,"/",model_wc,"/V_1_5_",yr,"_",model_wc,"_",model_wc_rcp,"_",climvar,".nc4?var=",climvar,"&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-01T12%3A00%3A00Z&time_end=",enddate,"T12%3A00%3A00Z&accept=csv_file",sep ="") 
      
      # http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/gcm/rcp85/inmcm4/V_1_5_2099_inmcm4_rcp85_soil_water_monthly.nc4?var=soil_water&latitude=45&longitude=-111&time_start=2099-01-16T05%3A14%3A31.916Z&time_end=2099-12-17T00%3A34%3A14.059Z&accept=csv_file
      
      holder <-data.frame(fread(data_url, 
                                verbose=FALSE, 
                                showProgress = FALSE,
      ))#temporary holder for subsets downloaded from cloud 
      
      mydat<-rbind(mydat, holder)#file that grows by adding each chunk downloaded from cloud 
      
      date <- mydat$time # dates don't download correctly without this
      
    }#end loop across years for rcp 8.5
    
    mydat2<-cbind(mydat2,mydat[,4])#append just the water balance data from each downloaded chunk 
    
    mydat<-NULL#reset this variable so it can accommodate a new column name given the new water balance variable it's extracting at each loop cycle 
    
  }#end loop across climate variables for 8.5
  
  
  
  mydat3<-cbind(date,holder[,2:3],mydat2)#join the data with metadata including date, lat, long 
  
  head(mydat3) 

  # ----------
  # Create loop for downloading the data rcp 4.5
  # ----------

  #initialize some variables to hold data

  holder <- NULL

  mydat <- NULL

  mydat2 <- NULL

  mydat4 <- NULL
  
  
  
  #This is the loop that runs across all the climate variables you want to download 
  
  #"agdd", fix this - return when agdd is back on the website

  for(climvar in c("soil_water", "runoff", "rain", "accumswe", "PET", "Deficit", "AET")){ #start loop for RCP 4.5



    print(paste("downloading",climvar, "daily", model_bc, model_bc_rcp))#this is a progress update as the loop runs


    for(yr in c(start:end)){# for each climate variable loop across the year.  allows users to select range if not interested in all data


      #Specify URL where are stored on cloud
      #leap years are written differently in the future data
      #ifelse for future data

      leap <- lubridate::leap_year(yr)
      if(leap == TRUE){
        enddate <- paste(yr,"12-31", sep = "-")
      } else {
        enddate <- paste(yr + 1, "01-01", sep = "-")
      }

      data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/gcm/",model_bc_rcp,"/",model_bc,"/V_1_5_",yr,"_",model_bc,"_",model_bc_rcp,"_", climvar, ".nc4?var=", climvar, "&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-01T12%3A00%3A00Z&time_end=",enddate,"T12%3A00%3A00Z&accept=csv_file",sep ="")

      holder <-data.frame(fread(data_url,
                                verbose=FALSE,
                                showProgress = FALSE,
      )) #temporary holder for subsets downloaded from cloud

      mydat<-rbind(mydat, holder) #file that grows by adding each chunk downloaded from cloud


      date <- mydat$time

    }#end loop across years for 4.5

    mydat2<-cbind(mydat2,mydat[,4])#append just the water balance data from each downloaded chunk

    mydat<-NULL#reset this variable so it can accommodate a new column name given the new water balance variable it's extracting at each loop cycle

  }#end loop across climate variables for 4.5

  mydat4<-cbind(mydat3,mydat2) #join the data with metadata including date, lat, long
  
  # "agdd_daily_wc", "agdd_daily_bc", - fix this, return when agdd is back on the website
  # make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)
  # 
  colnames(mydat4)[]<-c("date", "lat","lon","soil_water_daily_wc", "runoff_daily_wc", "rain_daily_wc",  "accumswe_daily_wc", "pet_daily_wc", "deficit_daily_wc", "aet_daily_wc", "soil_water_daily_bc", "runoff_daily_bc", "rain_daily_bc",  "accumswe_daily_bc", "pet_daily_bc", "deficit_daily_bc", "aet_daily_bc")

  
  # "agdd_daily_wc", fix this - not a part of the website 8-20-20
  ## make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)
  
  
# columns are multiplied by 10, need to fix this before writing csv
  
  divide.by.10 <- function(x, na.rm = FALSE) {
    x/10
  }
  
  mydat5 <- mydat4 %>% 
    mutate(lat = as.character(lat),
           lon = as.character(lon)) %>% 
    mutate_if(is.numeric, divide.by.10)
  
  write_csv(mydat5, here::here(site,
                               "raw_data",
                               paste("lat",lat,"lon",lon, sep = "_"),
                               "future_daily", 
                               paste(site,model_wc,model_wc_rcp, model_bc,model_bc_rcp, "daily_future.csv", sep = "_"))) #default is a space, sep = "" removes space

beep(10)

# here::here allows you to choose a folder in your working directory to save to
# here::here("folder I want to save to (this has to be made in your working directory before you save it)", "subfolder", "subfolder ( subfolders can be as numerous you want them to be, always separated by a comma)", "last thing in the list is the name of the object I'm saving")
```


### Monthly Future mixed models

```{r}

# download best case/worst case climate models, into one csv specified rcps only

start = start_future 

end = end_future

#initialize some variables to hold data 

holder <- NULL 

mydat <- NULL 

mydat2 <- NULL 

mydat3 <- NULL 



#This is the loop that runs across all the climate variables you want to download 

 # "agdd_monthly", fix this - return once agdd is back on the website
# make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)
  for(climvar in c("soil_water_monthly", "runoff_monthly", "rain_monthly","accumswe_monthly", "PET_monthly", "Deficit_monthly", "AET_monthly")){ 
  
  
  
  print(paste("downloading",climvar, model_wc, model_wc_rcp))#this is a progress update as the loop runs 
  
  
  
  #the url uses two variable names for same variable in monthly download string like "PET_monthly" for long name and "pet" for short name 
  
  #the short name is sometimes a different case than the long variable name, so make it lowercase so that 
  
  #url string will work with the format of data stored on the cloud 
  
  varshort<-unlist(strsplit(climvar,"_m"));varshort 
  
  vshort<-varshort[[1]];vshort 
  
  if(vshort=="PET"){ 
    
    vshort<-"pet" 
    
  };vshort 
  
  
  if(vshort=="Deficit"){#change case of short name from upper to lower case 
    
    vshort<-"deficit" 
    
  };vshort 
  
  
  if(vshort=="AET"){ 
    
    vshort<-"aet" 
    
  };vshort 
  
  
  for(yr in c(start:end)){# for each climate variable loop across the year.  allows users to select range if not interested in all data 
    
    
    #Specify URL where are stored on cloud 
    
leap <- lubridate::leap_year(yr) 
#leap years act differently with this monthly future data
#need to correct for that
    if(leap == TRUE){
      day <- 16
    } else {
      day <- 17
    }
    
    data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/gcm/",model_wc_rcp,"/",model_wc,"/V_1_5_",yr,"_",model_wc,"_",model_wc_rcp,"_",climvar,".nc4?var=",vshort,"&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-16T05%3A14%3A31.916Z&time_end=",yr,"-12-",day,"T00%3A34%3A14.059Z&accept=csv_file",sep ="") 
    
    
    
    holder <-data.frame(fread(data_url, 
                              verbose=FALSE, 
                              showProgress = FALSE,
    ))#temporary holder for subsets downloaded from cloud 
    
    mydat<-rbind(mydat, holder)#file that grows by adding each chunk downloaded from cloud 
    
    date <- mydat$time
    
    }#end loop across years for rcp 8.5
  
  mydat2<-cbind(mydat2,mydat[,4])#append just the water balance data from each downloaded chunk 
  
  mydat<-NULL#reset this variable so it can accodomate a new column name given the new water balance variable it's extracting at each loop cycle 
  
}#end loop across climate variables for rcp 8.5
  


mydat3<-cbind(date,holder[,2:3],mydat2)#join the data with metadat including date, lat, long 

head(mydat3) 


# ----------
# Create loop for downloading the data rcp 4.5
# ----------

#initialize some variables to hold data 

holder <- NULL 

mydat <- NULL 

mydat2 <- NULL 

mydat4 <- NULL 



#This is the loop that runs across all the climate variables you want to download 
# "agdd_monthly", fix this - removed because it was removed from website
# make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)
for(climvar in c("soil_water_monthly", "runoff_monthly", "rain_monthly", "accumswe_monthly", "PET_monthly", "Deficit_monthly", "AET_monthly")){ 
  
  
  
  print(paste("downloading",climvar, model_bc, model_bc_rcp))#this is a progress update as the loop runs 
  
  
  
  #the url uses two variable names for same variable in monthly download string like "PET_monthly" for long name and "pet" for short name 
  
  #the short name is sometimes a different case than the long variable name, so make it lowercase so that 
  
  #url string will work with the format of data stored on the cloud 
  
  varshort<-unlist(strsplit(climvar,"_m"));varshort 
  
  vshort<-varshort[[1]];vshort 
  
  if(vshort=="PET"){ 
    
    vshort<-"pet" 
    
  };vshort 
  
  
  
  if(vshort=="Deficit"){#change case of short name from upper to lower case 
    
    vshort<-"deficit" 
    
  };vshort 
  
  
  
  
  
  if(vshort=="AET"){ 
    
    vshort<-"aet" 
    
  };vshort 
  
  
  
  for(yr in c(start:end)){# for each climate variable loop across the year.  allows users to select range if not interested in all data 
    
    
    #Specify URL where are stored on cloud 
    
    leap <- lubridate::leap_year(yr) 
    if(leap == TRUE){
      day <- 16
    } else {
      day <- 17
    }
    
    data_url<-paste("http://www.yellowstone.solutions/thredds/ncss/daily_or_monthly/gcm/",model_bc_rcp,"/",model_bc,"/V_1_5_",yr,"_",model_bc,"_",model_bc_rcp,"_",climvar,".nc4?var=",vshort,"&latitude=",lat,"&longitude=",lon,"&time_start=",yr,"-01-16T05%3A14%3A31.916Z&time_end=",yr,"-12-",day,"T00%3A34%3A14.059Z&accept=csv_file",sep ="") 
    
    
    
    holder <-data.frame(fread(data_url, 
                              verbose=FALSE, 
                              showProgress = FALSE,
    ))#temporary holder for subsets downloaded from cloud 
    
    mydat<-rbind(mydat, holder)#file that grows by adding each chunk downloaded from cloud 
    
    date <- mydat$time
    
  }#end loop across years 
  
  mydat2<-cbind(mydat2,mydat[,4])#append just the water balance data from each downloaded chunk 
  
  mydat<-NULL#reset this variable so it can accommodate a new column name given the new water balance variable it's extracting at each loop cycle 
  


mydat4<-cbind(mydat3,mydat2) #join the data with metadata including date, lat, long
# "agdd_monthly_85", fix this - was removed because agdd was removed from website
## make sure the order matches where it was placed in the download order (i.e. if it downloads after rain, it should be placed here in the colnames after rain)
colnames(mydat4)[]<-c("date", "lat","lon","soil_water_monthly_wc", "runoff_monthly_wc", "rain_monthly_wc",  "accumswe_monthly_wc", "pet_monthly_wc", "deficit_monthly_wc", "aet_monthly_wc", "soil_water_monthly_bc", "runoff_monthly_bc", "rain_monthly_bc",  "accumswe_monthly_bc", "pet_monthly_bc", "deficit_monthly_bc", "aet_monthly_bc") 

divide.by.10 <- function(x, na.rm = FALSE) {
  x/10
}

mydat5 <- mydat4 %>% 
  mutate(lat = as.character(lat),
         lon = as.character(lon)) %>% 
  mutate_if(is.numeric, divide.by.10)


write_csv(mydat5, here::here(site, #requires project folder
                             "raw_data",
                             paste("lat",lat,"lon",lon, sep = "_"),
                             "future_monthly",
                             paste(site, model_wc, model_wc_rcp, model_bc, model_bc_rcp, "monthly_future.csv", sep = "_"))) #default is a space, sep = "" removes space

}# end loop across different climate projections

beep(10)

# here::here allows you to choose a folder in your working directory to save to
# here::here("folder I want to save to (this has to be made in your working directory before you save it)", "subfolder", "subfolder (subfolders can be as numerous you want them to be, always separated by a commma)", "last thing in the list is the name of the object I'm saving")


```

